---
title: "Pokemon Kaggle Analysis"
output: html_notebook
---


Import the data set

```{r}
library(data.table)
library(ggplot2)
```


```{r}
pokemon <- fread("./inputs/Pokemon.csv")
```


```{r}
colnames(pokemon)
print(pokemon, nrows=10)
```

We'll define these variables as:
$Total = HP + Attack + Defense + Sp. Atk + Sp. Def + Speed = Total Offense + Total Defense$
$Total Offense = Attack + Sp. Atk + Speed$
$Total Defense = Defense + Sp. Def + HP$

```{r}
pokemon <- pokemon[, "Total Offense":=Attack + `Sp. Atk` + `Speed`]
pokemon <- pokemon[, "Total Defense":=Defense + `Sp. Def` + `HP`]
```

# Exploration

Questions
1. how do I plot this as a barplot? Why aren't the columns in correct order?

**First: Lets get a top-level understanding of the distributions of various categories

```{r}
# distribution of pokemon counts
pokemon[,.N,by=`Type 1`][order(-N)]
ggplot(pokemon[,.N,by=`Type 1`][order(-N)], aes(x=`Type 1`, y=N)) + geom_bar(stat="identity")
```


```{r}
# average total power by type 1
pokemon[, .("Avg Total"=mean(Total), "Avg Total Offense"=mean(`Total Offense`), "Avg Total Defense"=mean(`Total Defense`)), by = `Type 1`][order(-`Avg Total`)]

# average total power by generation
pokemon[, .("Avg Total"=mean(Total), "Avg Total Offense"=mean(`Total Offense`), "Avg Total Defense"=mean(`Total Defense`)), by = Generation][order(-`Avg Total`)]

# what's the count of legendaries by generation
pokemon[Legendary==TRUE, .N, by = Generation]
```

> Dragon type Pokemon has the highest average type 1 value whereas Bug type 1 Pokemon have the lowest type 1 value on average.

```{r}
# average total among non legendaries
pokemon[Legendary==FALSE, 
        .("Avg Total"=mean(Total), "Avg Total Offense"=mean(`Total Offense`), "Avg Total Defense"=mean(`Total Defense`)), 
        by = `Type 1`][order(-`Avg Total`)]
```

> Dragon remains on top when when we exclude `Legendaries`

```{r}
# distribution of Total
ggplot(pokemon[,.(`Type 1`, Total)], aes(x=`Type 1`, y=Total)) + geom_boxplot()
```

